Relating two Abstractions for Formal Array Languages

### 2014-04-11 intromoa.tex
squeezing out text
I have edited as much as I can. Defs are essential.
If you can find anything else please go ahead and edit.
I know it is getting late for you and we need to get this in.
Let me know when you submit.
Lenore

MoA and the Psi Calculus, as a theory,  have  also been used to describe and verify 
hardware\cite{hardy1,hardy2,fftharry}, describe the partitioning and
scheduling of array operations (and their costs)\cite{mdst93,lee95,aachen96,coffin94,ll2,mcmahon95,haleh98}, 
and the formulation of parallel and distributed processing\cite{Mul03}. 
MoA and $\psi$-Calculus, can be used to abstract complex processor memory layouts 
for tensor/array expressions, reduce these expressions first to a semantic 
normal form (Denotational Normal Form-DNF) then to an Operational Normal Form 
(ONF). The ONF describes how to {\em build} the code using 
{\em start, stop, stride}, and {\em count}: a universal machine abstraction.
This is possible by {\em abstractly} lifting the dimension of an expression to include all aspects
of a processor, memory, communication topology. For example, to map a 2-d array expression to
n  processors we would lift the dimenion to 3-d. 
%
%
%
%\bibliographystyle{plain}
%\bibliography{paper1,hpf,workshop}
\subsection*{Elements of the theory}
\subsection*{\label{indshp} Indexing and Shapes}

@@

Most of the common array manipulation operations found in languages like Fortran90, Matlab, ZPL, etc.,
 can be defined from $\psi$ and a few elementary vector operations.

	We now introduce notation to permit us to define $\psi$ formally and to develop the
{\em Psi Correspondence Theorem}~\cite{muljen94}, which is central to the effective exploitation of MoA in array
computations. We will use $A, B, ...$ to denote an array of numbers (integer, real, or complex).
An array's dimensionality will be denoted by $d_A$ and will be assumed to 
be $n$ if not specified.

@@

	The precise mapping of $A$ to $a$ is determined by a one-to-one 
ordering function, $gamma$.  Although the choice of ordering is arbitrary, it is 
essential in the following that a specific one be assumed. By convention 
we assume the items of $A$ are placed in $a$ according to the lexicographic 
ordering of the indices of $A$. This is often referred to 
as $row \; major \; ordering$. 
Many programming languages lay out the items of multidimensional
arrays in memory in a contiguous segment using this ordering. 
Fortran uses the ordering corresponding to a transposed array in which the 
axes are reversed {\em column major}. Scalars are introduced as arrays with an empty shape vector.

@@

\hspace{.25cm}\= $u \;\cat\; v$ \hspace{1cm} \=		\=catentation of vectors $u$ and $v$ \\
\>$u$ + $v$\>		\>itemwise vector addition assuming $t_u = t_v$ \\
\>$u \times v$ \>	\>	itemwise vector multiplication \\
\>$n$ + $u$, $u$ + $n$ \>	addition of a scalar to each item of a vector \\
\>$n \times u$, $u \times n$ \>   multiplication of each item of a vector by a scalar \\
\>$\iota \; n$ \>\>  the vector of the first n integers starting from $0$ \\
%\footnote{$\iota \vec n$ is also defined and  produces an array of 
%cartesian coordinates~\cite{mul88}.}\\
\>$\pi \; v$ \> \>a scalar which is the product of the components of v \\
\>$k \;\take \;u$ \>\>	when $k \geq 0$ the vector of the first $k$ items of $u$,(called {\em take}) \\
\> \> \>and when $k<0$ the vector of the last $k$ items of $u$ \\
\>$k \;\drop \;u$ \>  \> when $k \geq 0$ the vector of $t_u - k$ last items of $u$,(called{ \em drop}) \\
\> \> \>and when $k<0$ the vector of the first $t_u - |k|$ items of $u$ \\
\end{tabbing}

>>>>

MoA and the Psi Calculus, as a theory,  have  also been used to describe and verify 
hardware\cite{hardy1,hardy2,fftharry}, describe the partitioning and
scheduling of array operations (and their costs)\cite{mdst93,lee95,aachen96,coffin94,ll2,mcmahon95,haleh98}, 
and the formulation of parallel and distributed processing\cite{Mul03}. 
MoA and $\psi$-Calculus, can be used to abstract complex processor memory layouts 
for tensor/array expressions, reduce these expressions first to a semantic 
normal form (Denotational Normal Form-DNF) then to an Operational Normal Form 
(ONF). 
%
%
%
%\bibliographystyle{plain}
%\bibliography{paper1,hpf,workshop}
\subsection*{Elements of the theory}
\subsection*{\label{indshp} Indexing and Shapes}

@@

	We now introduce notation to permit us to define $\psi$ formally. We will use $A, B, ...$ to denote an array of numbers (integer, real, or complex).
An array's dimensionality will be denoted by $d_A$ and will be assumed to 
be $n$ if not specified.

@@

	The precise mapping of $A$ to $a$ is determined by a one-to-one 
ordering function, $gamma$
\footnote{There are a family of gamma functions: 
$\gamma_{row}, \gamma_{column}, \gamma_{regular sparse}$, etc. In this case we assume row major ordering.}.  
Scalars are introduced as arrays with an empty shape vector.

@@

\hspace{.25cm}\= $u \;\cat\; v$ \hspace{1cm} \=		\=catentation of vectors $u$ and $v$ \\
\>$u$ + $v$\>		\>itemwise vector addition assuming $t_u = t_v$ \\
\>$u \times v$ \>	\>	itemwise vector multiplication \\
\>$n$ + $u$, $u$ + $n$ \>	addition of a scalar to each item of a vector \\
\>$n \times u$, $u \times n$ \>   multiplication of each item of a vector by a scalar \\
\>$\iota \; n$ \>\>  the vector of the first n integers starting from $0$ \\
%\footnote{$\iota \vec n$ is also defined and  produces an array of 
%cartesian coordinates~\cite{mul88}.}\\
\>$\pi \; v$ \> \>a scalar which is the product of the components of v \\
\>$k \;\take \;u$ \>\>	when $k \geq 0$ the vector of the first $k$ items of $u$) \\
\> \> \>and when $k<0$ the vector of the last $k$ items of $u$ \\
\>$k \;\drop \;u$ \>  \> when $k \geq 0$ the vector of $t_u - k$ last items of $u$) \\
\> \> \>and when $k<0$ the vector of the first $t_u - |k|$ items of $u$ \\
\end{tabbing}
